Efficiently Handling Guards in the low level design of RPython's tracing JIT

### 2012-07-20 talk/vmil2012/paper.tex
add the trace of the example as a figure
import example program
add an example figure showing how an operation and a following guard are compiled with and without merging
mention the figure in the text

@@

\section{Guards in the Backend}
\label{sec:Guards in the Backend}

After optimization the resulting trace is handed to the backend to be compiled
to machine code. The compilation phase consists of two passes over the lists of
instructions, a backwards pass to calculate live ranges of IR-level variables
and a forward one to emit the instructions. During the forward pass IR-level
variables are assigned to registers and stack locations by the register
allocator according to the requirements of the to be emitted instructions.
Eviction/spilling is performed based on the live range information collected in
the first pass. Each IR instruction is transformed into one or more machine
level instructions that implement the required semantics, operations withouth
side effects whose result is not used are not emitted. Guards instructions are
transformed into fast checks at the machine code level that verify the
corresponding condition.  In cases the value being checked by the guard is not
used anywhere else the guard and the operation producing the value can merged,
reducing even more the overhead of the guard. \bivab{example for this}

>>>>

\begin{figure}
    \input{figures/example.tex}
    \caption{Example Program}
    \label{fig:trace-log}
\end{figure}

@@

\begin{figure}
    \input{figures/log.tex}
    \caption{Optimized trace}
    \label{fig:trace-log}
\end{figure}
\section{Guards in the Backend}
\label{sec:Guards in the Backend}

After optimization the resulting trace is handed to the backend to be compiled
to machine code. The compilation phase consists of two passes over the lists of
instructions, a backwards pass to calculate live ranges of IR-level variables
and a forward one to emit the instructions. During the forward pass IR-level
variables are assigned to registers and stack locations by the register
allocator according to the requirements of the to be emitted instructions.
Eviction/spilling is performed based on the live range information collected in
the first pass. Each IR instruction is transformed into one or more machine
level instructions that implement the required semantics, operations withouth
side effects whose result is not used are not emitted. Guards instructions are
transformed into fast checks at the machine code level that verify the
corresponding condition.  In cases the value being checked by the guard is not
used anywhere else the guard and the operation producing the value can merged,
reducing even more the overhead of the guard. Figure \ref{fig:trace-compiled}
shows how an \texttt{int\_eq} operation followed by a guard that checks the
result of the operation are compiled to pseudo-assembler if the operation and
the guard are compiled separeated or if they are merged.

\bivab{Figure needs better formatting}
\begin{figure}[ht]
  \noindent
  \centering
  \begin{minipage}{1\columnwidth}
    \begin{lstlisting}
    i8 = int_eq(i6, 1)
    guard_false(i8) [i6, i1, i0]
    \end{lstlisting}
  \end{minipage}
  \begin{minipage}{.40\columnwidth}
    \begin{lstlisting}
CMP r6, #1
MOVEQ r8, #1
MOVNE r8, #0
CMP r8, #0
BEQ <bailout>
    \end{lstlisting}
  \end{minipage}
  \hfill
  \begin{minipage}{.40\columnwidth}
    \begin{lstlisting}
CMP r6, #1
BNE <bailout>
...
...
...
    \end{lstlisting}
  \end{minipage}
  \caption{Separated and merged compilation of operations and guards}
  \label{fig:trace-compiled}
\end{figure}
